State of the Practice for Lattice Boltzmann Method Software

### 2022-02-28 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/LatticeBoltz_SOP.tex
Completed first draft of section 9.2 (Circumventing the Oracle Problem and Addressing Reproducibility)

	\item[P\refstepcounter{pnum}\thepnum \label{P_Correctness}:]
	\textbf{Ensuring Correctness} Difficulties with ensuring correctness were
	noted by several developers. They alluded to difficulty with testing the
	correctness of large numbers of features, and challenges with automated
	automated testing. The TCLB developer commented that the amount of testing
	data needed was sometimes problematic, since free testing services do not
	offer adequate facilities for large amounts of data, which means in-house
	testing solutions are needed.  Other research software domains point to the
	following problems with testing: i) \citep{PintoEtAl2018} mention the
	problem of insufficient testing; ii) the survey of \citep{HannayEtAl2009}
	shows that more developers think testing is important than the number that
	believe they have a sufficient understanding of testing concepts; and, iii)
	\citep{HannayEtAl2009, KanewalaAndBieman2013, KellyEtAl2011, WieseEtAl2019}
	point to the oracle problem, which occurs in research software when we do
	not have a means to judge the correctness of the calculated solutions.  The
	LBM experience seems to overlap with i and ii, but the LBM developers did
	not allude to problem iii (the oracle problem) in our conversations.  As
	discussed in Section~\ref{SecReproducibility}, the oracle problem likely did
	not come up because the LBM developers we interviewed have strategies for
	verifying the reproducibility of their work.

@@

The top LBM developers show how modularization can be used to design research
software for future change.  Although the advice to modularize research software
to handle complexity is common \citep{WilsonEtAl2014, StewartEtAl2017,
Storer2017}, specific guidelines on how to divide the software into modules is
less prevalent.  Not every decomposition is a good design for supporting change,
as shown by \citet{Parns1972a}.  A design with low cohesion and
high coupling \citep[p.\ 48]{GhezziEtAl2003} will make change difficult.
Especially in research software, where change is inevitable, designers need to
produce a modularization that supports change.

LBM developers highlighted in our interviews cases where their modularizations
anticipated future changes.  The developer of pyLBM mentioned that the
geometries and models of their system had been ``decoupled'', using abstraction
and modularization of the source code, to make it ``very easy to add [new]
features''.  The pyLBM design allows for independent changes to the geometry and
the model.  We also learned that the package pyLBM redeveloped data structures
to ease future changes. The developer of TCLB noted that their software package
was designed to allow for the addition of some LBM features, but changes to
major aspects of the system would be difficult. For example, ``implementing a
new model will be an easy contribution'', but changes to the ``Cartesian mesh
... will be a nightmare''.  The design of TCLB highlights that not every
conceivable change needs to be supported, only the likely changes.  Likely
changes.

As the LBM developers illustrate, design for changes is accomplished by first
identifying likely changes, either implicitly or explicitly, and second by
hiding each likely change behind a well-defined module interface.  Although it
is unclear whether the developers are aware of Parnas's work, the approach
mirrors his recommendations \citep{Parns1972a}.

\subsection{Circumventing the Oracle Problem to Verify Reproducibility}
\label{SecReproducibility}

Software packages were not quantitatively measured for reproducibility. Instead,
developers were asked if they had concerns that their computational results
would not be reproducible in the future, and if they had taken any steps to
ensure reproducibility.  The developers all agreed that reproducibility is
important, but they differ in how far they go to support reproducibility.
ESPResSo does the most to support reproducibility.  For instance, they compare
the results of their methods against manually calculated results. These
comparison tests are automatically run for all source code changes. Continuous
integration is used, so these tests are run when a pull request is opened on
GitHub. Even once these tests are complete, a peer review process is undertaken
before changes are fully committed to the appropriate branch. The results for
all of the LBM schemes on the software package development branch are also
frequently compared for correctness, ensuring that the system output reflects
the expected output.

For other teams, the priority is verifying the mathematical foundations of the
models, but the output of the software package is not compared to other output.
Depending on the package and how it outputs solutions, this may not be practical
or feasible. A correct output may not be exactly reproducible, as it may be
dependent on a probability distribution, so strictly comparing results may not
be appropriate. \wss{Ideas for dealing with reproducibility in these situations
will be explored in the recommendations section (like properties of the solution,
such as conservation of mass).}

Some software packages, like Palabos, provide validation benchmarks for their
core fundamental algorithmic ingredients \citep{latt2021palabos}. The stated
intent of these benchmarks is to showcase the validity and usefulness of the
package to stimulate the development of third-party extensions. The Palabos
package identifies as a development framework for modelling problems in various
CFD areas.  \wss{Add validation benchmarks as a ``best practice'' used by
Palabos - others could emmulate.  Goal of this exercise is for people in the
community to learn from each other, and potentially from our advice too.}

Consider peer review processes to assess contributions and ensure the tracking
of information. High ranked package ESPResSo uses a peer review process for
contributions.

\wss{From Peter: The developer of ESPResSo mentioned ``tests which compare
implementations against each other, like say if there's to solve us for
electrostatic switch rely on different principles and they've been written by
different people''.
The developer of TCLB also mentioned automatic testing. They discussed the idea
of using test cases, but concerns were raised with this, such as the required
computing power to run such tests frequently and where to store large amounts of
data. Also, ``constructing such cases would be very hard''.
The developer of OpenLB commented that they have different kinds of tests, such
as integration tests, and tests where they know the ``exact analytical solution''.
They also mentioned comparing computations and ``check[ing] the misfits''.}

@@

@@

@@

>>>>

	\item[P\refstepcounter{pnum}\thepnum \label{P_Correctness}:]
	\textbf{Ensuring Correctness} Difficulties with ensuring correctness were
	noted by several developers. They alluded to difficulty with testing the
	correctness of large numbers of features, and challenges with automated
	automated testing. The TCLB developer commented that the amount of testing
	data needed was sometimes problematic, since free testing services do not
	offer adequate facilities for large amounts of data, which means in-house
	testing solutions are needed.  Other research software domains point to the
	following problems with testing: i) \citep{PintoEtAl2018} mention the
	problem of insufficient testing; ii) the survey of \citep{HannayEtAl2009}
	shows that more developers think testing is important than the number that
	believe they have a sufficient understanding of testing concepts; and, iii)
	\citep{HannayEtAl2009, KanewalaAndBieman2013, KellyEtAl2011, WieseEtAl2019}
	point to the oracle problem, which occurs in research software when we do
	not have a means to judge the correctness of the calculated solutions.  The
	LBM experience seems to overlap with i and ii, but the LBM developers did
	not allude to problem iii (the oracle problem) in our conversations.  As
	discussed in Section~\ref{SecReproducibility}, the oracle problem likely did
	not come up because the LBM developers we interviewed have strategies for
	verifying their work.

@@

To address technical debt (\pref{P_TechDebt}), the top LBM developers show how
modularization can be used to design research software for future change.
Although the advice to modularize research software to handle complexity is
common \citep{WilsonEtAl2014, StewartEtAl2017, Storer2017}, specific guidelines
on how to divide the software into modules is less prevalent.  Not every
decomposition is a good design for supporting change, as shown by
\citet{Parnas1972a}.  A design with low cohesion and high coupling \citep[p.\
48]{GhezziEtAl2003} will make change difficult. Especially in research software,
where change is inevitable, designers need to produce a modularization that
supports change.

LBM developers highlighted in our interviews cases where their modularizations
anticipated future changes.  The developer of pyLBM mentioned that the
geometries and models of their system had been ``decoupled'', using abstraction
and modularization of the source code, to make it ``very easy to add [new]
features''.  The pyLBM design allows for independent changes to the geometry and
the model.  We also learned that the package pyLBM redeveloped data structures
to ease future changes. The developer of TCLB noted that their software package
was designed to allow for the addition of some LBM features, but changes to
major aspects of the system would be difficult. For example, ``implementing a
new model will be an easy contribution'', but changes to the ``Cartesian mesh
... will be a nightmare''.  The design of TCLB highlights that not every
conceivable change needs to be supported, only the likely changes.

As the LBM developers illustrate, design for changes is accomplished by first
identifying likely changes, either implicitly or explicitly, and second by
hiding each likely change behind a well-defined module interface.  Although it
is unclear whether the developers are aware of Parnas's work, the approach
mirrors his recommendations \citep{Parnas1972a}.

\subsection{Circumventing the Oracle Problem and Addressing Reproducibility}
\label{SecReproducibility}

The top LBM projects have adopted techniques for software verification that get
around the oracle problem (mentioned earlier in Section~\ref{P_Correctness}). In
the absence of a general test oracle, the following techniques have been adopted
by LBM developers:

\begin{itemize}
	\item Compare the calculated results against manually calculated results in
	the cases where manual solutions can be determined.  Although it is not
	always feasible to calculate results a priori, developers can find sub-sets
	of the general problem where answers can be determined.  In special cases
	that allow this, the results can be compared against exact analytical
	solutions, as done by OpenLB.
	\item Related to peer review, verify the mathematical foundations of the
	models to build confidence in the software's theoretical underpinnings.
	\item Compare the calculated results against pseudo-oracle solutions for
	verification benchmarks. The developer of ESPResSo mentioned the value of
	``tests which compare implementations against each other'' where the tests
	``rely on different principles and they've been written by different
	people''. \citet{latt2021palabos} shows several benchmark tests comparing
	Palabos solutions to other software solutions, such as solutions calculated
	using AnSys.
	\item Use peer review of the code for major changes and additions, as
	mentioned in Section~\ref{Sec_CompareMethodologies} for ESPResSo. 
\end{itemize}

\noindent Additional strategies for getting around the oracle problem are
provided in Section~\ref{Sec_Recommendations}.

Some LBM developers improve reproducibility via automating the verification
tests and making them part of a continuous integration process (mentioned in
Section~\ref{Sec_CompareTools}).  Providing unit tests improves reproducibility
because future developers will require a means to verify that they can obtain
the same computational results as today.  Unit tests are a good way to embed
self-diagnostics of reproducibility in the code \citep{BenureauAndRougier2017}.

@@

Borrow from SE for HEP paper.

@@

\subsection{Verifiability}

- further steps to address the oracle problem - borrow from SE for HEP paper.
- anything in Smith 2016?

@@

Repeatability - A correct output may not be exactly reproducible, as it may be
dependent on a probability distribution, so strictly comparing results may not
be appropriate. \wss{Ideas for dealing with reproducibility in these situations
will be explored in the recommendations section (like properties of the
solution, such as conservation of mass).  5R paper mentions repeatability -
design it so that the random number seed can be set.}
